---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<h1 align="center">FTM - Flexible Transfer Models</h1>
<p align="center"><i>Seamlessly transfer omics models between datasets with varying predictors.</i></p>
<div align="center">
  <a href="https://www.r-project.org/"><img src="https://img.shields.io/badge/r-%23276DC3.svg?style=for-the-badge&logo=r&logoColor=white" alt="R"/></a>
  <a href="https://www.baker.edu.au/"><img src="https://img.shields.io/badge/Baker-%23276DC3?style=for-the-badge" alt="Baker Institute"/></a>
  <br>
  <p align="center"><i>Love our work? Visit our <a href="https://metabolomics.baker.edu.au/">WebPortal</a>.</i></p>
</div>

# Overview of the FTM package

Often, with 'omics data, we have a model that we have trained on one dataset and we want to apply it to another dataset. However, the predictors in the new dataset may not be entirely the same as the predictors in the original dataset. This causes an issue, which can traditionally be solved through 2 options:

- Ignore the missing predictors (effectively setting them to 0)
- Recreate the model using a subset of variables that are present in both datasets

FTM overcomes this challenge by allowing the model to be evaluated in datasets without exact overlap in the predictor variables, without loss in generalisability, re-optimization, or access to individual-level data. This is achieved by storing intermediate matrices that allow beta coefficients to be recalculated using a subset of original predictors.


## Installation

You can install the development version of FTM from [GitHub](https://github.com/) with:

```r
if (!require("devtools")) install.packages("devtools")
if (!require("glmnet")) install.packages("glmnet")
devtools::install_github("BakerMetabolomics/FlexibleTransferModels", build_vignettes = TRUE)
```

Load the package. You'll need to have [glmnet](https://glmnet.stanford.edu/articles/glmnet.html) installed to use the FTM package.

```{r message=TRUE, warning=FALSE}
library(FTM)
```
## Example

This is a basic example which shows you how to convert an existing model to a FTM object and predict outcomes in new dataset:

```{r example}
# Set up the predictors (glmnet requires a matrix)
predictors <- as.matrix(mtcars[, c("hp", "wt", "cyl")])

# 1. Fitting a glmnet model to the mtcars dataset
glmnet_model <- glmnet::cv.glmnet(predictors, mtcars$am, family = "binomial", alpha = 0)

# 2. Create FTM object using createFTM
ftmglmnet_model <- createFTM(glmnet_model, x = predictors, y = mtcars$am)

# A quick look at the model
show(ftmglmnet_model)

# Extract the beta coefficients from a FTM object
coef(ftmglmnet_model)


# 3. Predict on "new" data
new_data <- mtcars[1:10, c("hp", "wt", "cyl")]

# Predict using the FTM from glmnet model
predictions <- predict(ftmglmnet_model, newdata = new_data, type = "response")
predictions

# 4. Predict on "new" data that has missing predictors
new_data_missing <- mtcars[1:10, c("hp", "wt")] # missing cyl

# Predict using the FTM from glmnet model
predictions_missing <- predict(ftmglmnet_model, newdata = new_data_missing, type = "response")
predictions_missing

```
## Documentation

For more detailed information and advanced usage, please refer to the package vignettes available within R:

```r
browseVignettes("FTM")
```

## Contributing

Contributions to the FTM package are welcome from the community. Bug reports, documentation enhancements, and feature requests can be filed on the [GitHub issues page](https://github.com/BakerMetabolomics/FlexibleTransferModels/issues).


