---
title: "Estimating performance when you have fewer variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimating performance when you have fewer variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
## load packages
library(glmnet)
library(FTM)
```

# Using FTMlm objects to estimate the model performance
FTMlm objects allow for the estimation of model performance (R2) for the full model and a model using a subset of variables. This can be done using the `rsq` function.

```{r}
# Load the mtcars dataset
data(mtcars)

# Fit a linear model to the mtcars dataset, predicting mpg from cyl, hp, and wt
lm_model <- lm(mpg ~ cyl + hp + wt, data = mtcars)

# Create FTM object using createFTM
ftmlm_model <- createFTM(lm_model) # same as using createFromLm

# Calculate the model performance
rsq(ftmlm_model)

# Check the linear model R2
summary(lm_model)$r.squared
```

Notice the ftmlm model reports the R2 value and that it matches the linear model it was created from. This R2 value was calculated using the parameters stored within the object.

# Estimating model performance when using a subset of variables

```{r}
# Calculate the model performance for a model using just cyl + wt
rsq(ftmlm_model, select = c("cyl", "wt"))

# Verify we get the same performance as a de novo model
summary(lm(mpg ~ cyl + wt, data = mtcars))$r.squared


# Calculate the model performance for a model using just hp
rsq(ftmlm_model, select = c("hp"))

# Verify we get the same performance as a de novo model
summary(lm(mpg ~ hp, data = mtcars))$r.squared
```

# Notes

## Ridge penalties

Estimating model performance with a ridge penalty causes the solution to not be exact. The estimate is often rather good, but may slightly under/over estimate the model performance. If this of concern, compare the model performance with the optimal `s` and with `s = 0`. If these are similar, consider using the estimate from the non-ridge penalty model.

## GLM models

Currently, the estimation of model performance hasn't been implemented for ftmglm models. Once we are happy with a solution, we will implement this feature.
