---
title: "Extracting coefficients"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting coefficients}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
## load packages
library(glmnet)
library(FTM)
```

```{r include=FALSE}
data(mtcars)

# Fitting a linear model to the mtcars dataset
lm_model <- lm(mpg ~ cyl + hp + wt, data = mtcars)

# Create FTM object using createFromLm
ftmlm_model <- createFromLm(lm_model)

# Create FTM object using createFTM
ftmlm_model <- createFTM(lm_model) # same as using createFromLm

# Fitting a glm to the mtcars dataset
glm_model <- glm(am ~ hp + wt + cyl, data = mtcars, family = "binomial")

# Create FTM object using createFromGlm
ftmglm_model <- createFromGlm(glm_model)

# Create FTM object using createFTM
ftmglm_model <- createFTM(glm_model)

# set the predictors
predictors <- as.matrix(mtcars[, c("hp", "wt", "cyl")])

# Fitting a glmnet to the mtcars dataset
glmnet_model <- glmnet::cv.glmnet(predictors, mtcars$am, family = "binomial")

# Create FTM object using createFromGlmnet
ftmglmnet_model <- createFromGlmnet(glmnet_model, x = predictors, y = mtcars$am)

# Create FTM object using createFTM
ftmglmnet_model <- createFTM(glmnet_model, x = predictors, y = mtcars$am)
```

# Extract beta coefficient from FTM object

FTM objects allow you to view beta coefficients for all variables or for a subset of variables. The `coef` function can be used to extract coefficients from a fitted FTM object. The `select` argument can be used to extract coefficients for specific variables (accepts variable names or integers).

```{r}
# Assuming ftmglm_model, ftmlm_model or ftmglmnet_model are pre-fitted model objects (see Create FTM Object vignette)

# Extract beta coefficients for all variables from a ftmlm model
coef(ftmlm_model)

# Extract beta coefficients for a reweighted model using a subset of variables from a ftmlm model
coef(ftmlm_model, select = c("cyl", "hp"))



# This works similarly for a ftmglm model
coef(ftmglm_model)

# Similarly, you can extract beta coefficients after reweighting for a subset of variables from a ftmglm model
coef(ftmglm_model, select = c("cyl", "hp"))



# This also works for FTM objects made from glmnet models
coef(ftmglmnet_model)

# Finally, you can extract beta coefficients after reweighting for a subset of variables from a ftmglmnet model
coef(ftmglmnet_model, select = c("cyl", "hp"))
```

# Manually subset of FTM object

Although a FTM object contains information about all variables, it is possible to manually `trim` a FTM model to a subset of variables.

```{r}
# Assuming ftmglm_model, ftmlm_model or ftmglmnet_model are pre-fitted model objects (see Create FTM Object vignette)

# Subset a ftmlm model, so it only contains a subset of variables
subset(ftmlm_model, subset = c("cyl", "hp"))



# Similarly, you can subset a ftmglm model
subset(ftmglm_model, subset = c("cyl", "hp"))



# This also works for FTM objects made from glmnet models
subset(ftmglmnet_model, subset = c("cyl", "hp"))
```
