---
title: "Predicting in new datasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Predicting in new datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r message=FALSE, warning=FALSE, include=FALSE}
## load packages
library(glmnet)
library(FTM)
```

```{r include=FALSE}
data(mtcars)

# Fitting a linear model to the mtcars dataset
lm_model <- lm(mpg ~ cyl + hp + wt, data = mtcars)

# Create FTM object using createFromLm
ftmlm_model <- createFromLm(lm_model)

# Create FTM object using createFTM
ftmlm_model <- createFTM(lm_model) # same as using createFromLm

# Fitting a glm to the mtcars dataset
glm_model <- glm(am ~ hp + wt + cyl, data = mtcars, family = "binomial")

# Create FTM object using createFromGlm
ftmglm_model <- createFromGlm(glm_model)

# Create FTM object using createFTM
ftmglm_model <- createFTM(glm_model)

# set the predictors
predictors <- as.matrix(mtcars[, c("hp", "wt", "cyl")])

# Fitting a glmnet to the mtcars dataset
glmnet_model <- glmnet::cv.glmnet(predictors, mtcars$am, family = "binomial")

# Create FTM object using createFromGlmnet
ftmglmnet_model <- createFromGlmnet(glmnet_model, x = predictors, y = mtcars$am)

# Create FTM object using createFTM
ftmglmnet_model <- createFTM(glmnet_model, x = predictors, y = mtcars$am)
```

# Using FTM objects to predict outcomes in new datasets
FTM objects allow the prediction of outcomes using the `predict` function, just like the original model.

```{r}
# Create a "new" data
new_data <- mtcars[1:10, c("hp", "wt", "cyl")]

# predict using the FTM from lm model
predict(ftmlm_model, newdata = new_data)

# predict using the FTM from glm model
predict(ftmglm_model, newdata = new_data)

# predict using the FTM from glmnet model
predict(ftmglmnet_model, newdata = new_data)
```

# FTM objects allow predictions even with missing variables
The major benefit of FTM objects is that they allow predictions even when the new dataset has missing variables. This is because of the flexible structure of the FTM that is achieved by reweighting of the beta coefficients.

```{r}
# Create a "new" data with missing variables (missing "cyl")
new_data <- mtcars[1:10, c("hp", "wt")]

# predict using the FTM from lm model
predict(ftmlm_model, newdata = new_data)

# predict using the FTM from glm model
predict(ftmglm_model, newdata = new_data)

# predict using the FTM from glmnet model
predict(ftmglmnet_model, newdata = new_data)
```
